\chapter{Introduction}
\section{Algebra of Complex Numbers}
We define complex numbers to be all the pairs of real numbers \((x,y)\) with following addition and multiplication:
\[(x_1,y_1) +' (x_2,y_2) = (x_1 + x_2 , y_1 + y_2)\]
\[ (x_1,y_1) \cdot' (x_2,y_2) = (x_1x_2 - y_1y_2, x_1y_2 + x_2y_1)\]
where \(+\) and \(\cdot\) are real addition and multiplication, respectively.
We denote the set of complex numbers with \(\Complex\).
It is easy to check that addition and multiplication defined above have the following:
\begin{enumerate}
        \item Addition and multiplication are commutitive:
              \[z + w = w + z \quad z,w \in \Complex\]
              \[ z\cdot w = w \cdot z \quad z,w \in \Complex\]

        \item Addition and multiplication are associative:
              \[(z + w) + u = z + (w + u) \quad z,w,u \in \Complex\]
              \[ (z \cdot w) \cdot u = z \cdot (w \cdot u) \quad z,w,u \in \Complex\]
        \item Addition and multiplication are distributive:
              \[(z + w) \cdot u = z \cdot u + w \cdot u \quad z,w,u \in \Complex\]
        \item Addition and multiplication have unique identity elements \(0 = (0,0)\) and \(1 = (1,0)\), respectively.
        \item Every complex number \(z\) has a unique addition inverse. Denoted by \(-z\).
        \item Every non-zero complex number \(z\) has a unique multiplication inverse. Denoted by \(z^{-1}\) or \(\dfrac{1}{z}\).
\end{enumerate}
Which means \(\Complex\) is a field. 

We can represent in many other forms. Two of the most commonly for \(z = (x,y)\) used are :
\[ z = x + iy \quad \text{where} \; i = (0,1)\]
\[ z = re^{i\theta} \quad r \geq 0 \,, \theta \in \Reals \]
\[ z = \begin{bmatrix}
                x & -y \\
                y & x
        \end{bmatrix} \]
It is easy to see that \(i^2 = -1\). We also define the functions \(\func{\Re}{z} = x\) and \(\func{\Im}{z} = y\). In the second representation \(r\) is the distance from origin and \(\theta\) is the angle between the positive real axis and the ray passing through \(z\). One can also view complex number as an extension of real numbers in which every polynomial has a root.

\begin{theorem}
      If \(K\) is field such that every odd degree polynomial has at a least a root and for all \(a \in K\), either \(x^2 = a\) or \(x^2 = -a\) has a root then it is sufficient to \textit{add} the root of \(x^2 = -1\) to \(K\) so that every polynomial has a root.
\end{theorem}

For every complex number \(z = x + iy\) there exists the mapping \(\bar{z}: \Complex \to \Complex\) where \(\bar{z} = x - iy\) and is called the \(\emph{conjugate}\) of \(z\).
\begin{proposition}
        The following properties are satisfied.
        \begin{enumerate}
                \item \(\overline{z + w} = \bar{z} + \bar{w}\).
                \item \(\overline{zw} = \bar{z}\bar{w}\).
                \item \(z = \bar{\bar{z}}\).
                \item \(z = \bar{z}\) if and only if \(z \in \mathbb{R}\).
                \item The following relations hold:
                      \[ \func{\Re}{z} = \dfrac{z + \bar{z}}{2} \quad,\quad \func{\Im}{z} = \dfrac{z - \bar{z}}{2i}\]
        \end{enumerate}
\end{proposition}
Another mapping is the \(\emph{norm}\) or \(\emph{modulus}\) function, \(\abs{z}: \Complex \to \Reals \) where \(\abs{z} = \sqrt{x^2 + y^2}\). Geometrically speaking the norm of \(z\) gives the distance of \(z\) form origin.
\begin{proposition}
        The following properties are satisfied.
        \begin{enumerate}
                \item \(\abs{z} \geq 0 \quad \forall z \in \Complex\) and especially \(\abs{z} = 0\) if and only if \(z = 0\)
                \item \(\abs{zw} = \abs{z}\abs{w}\).
                \item \(\abs{z}^2 = z\bar{z}\).
                \item \(\abs{z} = \abs{-z} = |\bar{z}|\)
                \item The following inequalities hold:
                      \[ -\abs{z} \leq \func{\Re}{z} \:,\: \func{\Im}{z} \leq \abs{z}\]
                \item Triangle inequality:
                      \[ \abs{z + w} \leq \abs{z} + \abs{w} \]
                      \[ \abs{\abs{z} - \abs{w}} \leq \abs{z- w}\]
        \end{enumerate}
\end{proposition}

The point at infinity, \(\infty\), is informally a point that is unboundedly far from the origin. The extended complex plane is defined as \(\Complex \cup \set{\infty}\). Every line passes through \(\infty\) but no half plane contains it.

\section{Riemann sphere}
\textbf{Riemann sphere} is a unit sphere centered at the origin complex plane. We can map every point (except \(\bracket{0,0,1}\)) on the sphere to a point in complex plane by the following bijective transformation.
\begin{equation*}
      \func{\phi}{z} = \bracket{x_1, x_2, x_3}= \bracket{\dfrac{z + \bar{z}}{1 + \abs{z}^2} , \dfrac{z - \bar{z}}{i \bracket{1 + \abs{z}^2}} , \dfrac{ \abs{z}^2 - 1}{\abs{z}^2 + 1} }
\end{equation*}
Geometrically, this transformation maps every point \(z\) to the interestion of the line connecting it to the \(\bracket{0,0,1}\) and the sphere. Furthermore, we can define \(\func{\phi{\infty}} = \bracket{0,0,1}\).

\section{Limits and continuity}

Let \(f : U \to \Complex\), \(\alpha\) be an adherent point of \(U\), and \(w\) be a complex number. Then 
\begin{equation*}
      w = \lim_{z \to \alpha} \func{f}{z} 
\end{equation*}
when 
\begin{equation*}
      \forall \epsilon > 0, \; \exists \delta > 0 \ \suchThat \ z \in U , \abs{z - \alpha} < \delta \implies \abs{\func{f}{z} - w} < \epsilon
\end{equation*}
By defining the \(\epsilon-\)neighbourhood of \(\infty\) to the 
\begin{equation*}
      \set<z>{\dfrac{1}{z} < \epsilon}
\end{equation*}
we can extend the defintion of limit to when \(\alpha\) or \(w\) are \(\infty\).
We say that \(f\) is continuous at \(\alpha\) if 
\begin{equation*}
      \lim_{z \to \alpha} \func{f}{z} = \func{f}{\alpha}
\end{equation*}

\begin{definition}
      A function \(f\) is said to be uniformly continuous if 
      \begin{equation*}
            \forall \epsilon > 0 , \; \exists \delta > 0 \ \suchThat \ z,w \in S , \abs{z - w} < \delta \implies \abs{\func{f}{z} - \func{f}{w}}
      \end{equation*}
\end{definition}

\section{Sequences and convergence}
A Sequence \(\set{z_n}_{n \in \Naturals}\) is said to be convergent to \(z\) if:
\begin{equation*} 
      \forall \epsilon > 0 , \; \exists N \ \suchThat \ n \geq N \implies \abs{z - z_n} < \epsilon
\end{equation*}
and it is denoted as 
\begin{equation*}
      z = \lim_{n \to \infty} z_n
\end{equation*}

A sequence \(\set{z_n}\) is a \textbf{Cauchy sequence} if 
\begin{equation*}
      \forall \epsilon, \; \exists N \ \suchThat \ n,m \geq N \implies \abs{z_m - z_n} < \epsilon
\end{equation*}
If we write \(z_n = x_n + iy_n\), since \(\abs{z_n - z_m} = \sqrt{(x_n - x_m)^2 + (y_n - y_m)^2}\) and \(\abs{x_n - x_m}\leq \abs{z_n - z_m} \), \( \abs{y_n - y_m} \leq \abs{z_n - z_m}\), we can conclude that \(\set{z_n}\) is Cauchy if and only if \(\set{x_n}\) and \(\set{y_n}\) are Cauchy sequences. Thus, since for real sequences a Cauchy sequence is convergent then, a complex Cauchy sequence is convergent as well.

Due to similarities between the definition given above and their real counterpart, most results dealing with limits can be easily extended to complex numbers.

\section{Compact sets}
\begin{definition}[Point of accumulation]
      point \(v\) is a \textbf{point of accumulation} for the sequence \(\set{z_n}\) if for given \(\epsilon > 0\) there exists infinitely many \(n\) such that 
      \begin{equation*}
            \abs{z_n - v} < \epsilon
      \end{equation*}
      Similarly, a point of accumulation of an infinite set \(S\) is a point \(v\) that for each open set \(U\) containing \(v\) there are infinitely many elements of \(S\).
\end{definition}

\begin{theorem}[Weierstrass-Bolzano theorem]
      If \(S\) is an infinite bounded set of real numbers, then \(S\) has a point of accumulation.
\end{theorem}

\begin{definition}
      A \textbf{compact} set \(S\) is a set that every sequence of its elements has a point of accumulation in \(S\). The following defintions are equivalent 
      \begin{enumerate}
            \item Every infinite subset of \(S\) has a point of accumulation in \(S\).
            \item Every sequence of elements of \(S\) has a convergent subsequent whose limit is in \(S\).
      \end{enumerate}
\end{definition}

\begin{theorem}
      A complex set is compact if and only if it is closed and bounded.
\end{theorem}

\begin{theorem}
      Let \(S_1 \supset S_2 \supset \dots\) be a sequence of non-empty closed subsets of a compact set \(S\). Then, the interestion of all \(S_n\) is not empty.
\end{theorem}

\begin{theorem}
      Let \(S\) be a compact set and \(f\) be continuous function on \(S\). Then the image of \(f\) is compact.
\end{theorem}

\begin{theorem}
      Let \(S\) be a compact set and \(f\) be continuous function on \(S\). Then  \(f\) is uniformly continuous.
\end{theorem}


\begin{definition}
      Let \(A,B\) be two sets of complex numbers. The \textbf{distance} between them is 
      \begin{equation*}
            \func{d}{A,B} = \min_{\substack{\alpha \in A \\ \beta \in B}} \abs{\alpha - \beta}
      \end{equation*}
\end{definition}

\begin{theorem}
      Let \(S\) be a closed set and let \(v\) be a complex number. There exists a point \(w \in S\) such that 
      \begin{equation*}
            \func{d}{S,\set{v}} = \abs{w - v}
      \end{equation*}
\end{theorem}

\begin{theorem}
      Let \(K\) be a compact set and let be \(S\) a closed set. Then, there are elements \(\alpha \in K\) and \(\beta \in S\) such that 
      \begin{equation*}
            \func{d}{K,S} = \abs{\alpha - \beta}
      \end{equation*}
\end{theorem}

\begin{theorem}
      Let \(K\) be compact. Let \(r\) be a real number greater than zero. There exists a finite number of discs of radius \(r\) whose union contains \(K\).
\end{theorem}

We say that a family of open set \(\set{U_i}\) covers a set \(S\) when for every \(z \in S\), \(z \in U_i\) for some \(i\) as well. A subcovering of \(S\) is a covering of \(S\) with a subfamily of \(\set{U_i}\). If that subfamily is finite we say that it is a finite subcovering of \(S\). \(S\) is \textbf{covering compact} if every open convering can reduce to a finite subcovering.

\begin{theorem}
      Let \(S\) be a set then \(S\) is sequentially comapct if and only if covering compact.
\end{theorem}

\section{Differentiability}
Let \(f: U \overset{open}{\subset} \Complex \to \Complex\) and \(z\) be a point in \(U\). \(f\) is said to be complex differentiable at \(z\) if the following limit exists.
\begin{equation*}
      \func{f'}{z} = \lim_{h \to 0} \dfrac{\func{f}{z + h} - \func{f}{z}}{h}
\end{equation*}
Equivalently, there exists a \(c \in \Complex\) such that 
    \begin{equation} 
        \func{f}{z_0 + h} - \func{f}{z_0} - hc = \littleO{h} 
\end{equation}
Differentiation rules are similar to real differentiation and are proved the same way. 

\(f\) is \textbf{holomorphic} if it is differentiable at every point of \(U\). Taking the limit along the real and imaginary axes and equating them gives the \textbf{Cauchy-Riemann equations}

\begin{equation*}
      \PDiff{u}{x} = \PDiff{v}{y} ,\quad \PDiff{u}{y} = - \PDiff{v}{x}
\end{equation*}

Assuming that \(u\) and \(v\) are twice differentiable, we can see that 
\begin{align*}
      \Delta u &= \dfrac{\PDiffOperator^2 u}{\PDiffOperator x^2} + \dfrac{\PDiffOperator^2 u}{\PDiffOperator y^2}\\
      &= \dfrac{\PDiffOperator^2 v}{\PDiffOperator x \PDiffOperator y} - \dfrac{\PDiffOperator v}{\PDiffOperator y \PDiffOperator x} = 0
\end{align*}
Which means that \(u,v\) satisfy the \textit{Laplace's equation} and thus are harmonic functions. 

\begin{theorem}
      If \(u,v\) and have continuous first order partial and satisfy the Cauchy-Riemann equations then \(\func{f}{z} = \func{u}{x,y} + \func{v}{x,y}\) is analytic with continuous derivative.
\end{theorem}

\section{Polynomials and rational functions}

\begin{theorem}
      If all zero's of a polynomial \(P\) lie in a half plane (convex polygon), the all zero's of the derivative \(P'\) lie in the same half plane (convex polygon).
\end{theorem}

A \textbf{ration function} \(\func{R}{z}\) is the quotient of two polynomials
\begin{equation*}
      \func{R}{z} = \dfrac{\func{P}{z}}{\func{Q}{z}} = \dfrac{a_m x^m + \dots + a_0}{b_n x^n + \dots + b_0}
\end{equation*}
We can assume that \(P\) and \(Q\) have no factor in common. In that case, the zeros of \(P\) are the zeros of \(R\) and the zeros of \(Q\) are called the \textit{poles} of \(R\). We can further define \(\func{R}{\infty}\) to be \(\func{S_R}{0}\) where \(\func{S_R}{z} = \func{R}{\frac{1}{z}}\). 
\begin{equation*}
      \func{S_R}{z} = z^{n-m} \dfrac{a_m + \dots + a_0 x^m}{b_n + \dots + b_0 x^n}
\end{equation*}
Then if \(n > m\) we say that \(R\) has roots at \(\infty\) with multiplicity of \(n - m\), if \(n < m\) then \(R\) has poles at \(\infty\) with multiplicity of \(m  - n\). If \(n = m\) then \(\func{R}{\infty} = \frac{a_m}{b_n}\). It is now easy to see that the number of roots and poles of \(R\) - including \(\infty\) - is the greater of \(m,n\) and it is called the \textit{degree} of \(R\). 

We can write \(R\) in the following form 
\begin{equation*}
      \func{R}{z} = \func{G}{z} + \func{H}{z}
\end{equation*}
where \(G\) is a polynomial with no constant term, by carrying out the division. Now let \(\beta_1 , \dots , \beta_k\) be distinct poles of \(R\). Then \(\func{S_j}{\xi} = \func{R}{\beta_j + \frac{1}{\xi}}\) is a rational function of \(\xi\) and it can be written as 
\begin{equation*}
      \func{S_j}{\xi} = \func{G_j}{\xi} + \func{H_j}{\xi} \implies \func{R}{z}= \func{G_j}{\dfrac{1}{z - \beta_j}} + \func{H_j}{\dfrac{1}{z - \beta_j}}
\end{equation*}
note that \(\func{G_j}{\frac{1}{z - \beta_j}}\) has pole(s) only at \(\beta_j\). Now consider 
\begin{equation*}
      \func{S}{z} = \func{R}{z} - \func{G}{z} - \sum_{j = 1}^k \func{G_j}{\dfrac{1}{z - \beta_j}}
\end{equation*}
It can have poles only at \(\beta_1 , \dots , \beta_k\) and \(\infty\). \(R\) and \(G_j\) have poles at \(z = \beta_j\), however their difference if finite. \(R\) and \(G\) have poles at \(\infty\) but their difference is finite as well. Hence \(S\) has no poles and thus must be a constant. Incorporating the constant into \(G\) allows us to write \(R\) as 
\begin{equation*}
      \func{R}{z} = \func{G}{z} + \sum_{j = 1}^k \func{G_j}{\dfrac{1}{z - \beta_j}}
\end{equation*}
which is its \textit{partial fraction decomposition}.

\section{Conformal maps}
A \textbf{conformal map} is a transformation that conserves angle and direction. For example every linear \(\Reals^2\) transformation of the form 
\begin{equation*}
      \begin{bmatrix}
            a & -b\\
            b & a
      \end{bmatrix} = \sqrt{a^2 + b^2} \begin{bmatrix}
            \cos \theta & -\sin \theta\\
            \sin \theta & \cos \theta
      \end{bmatrix} 
\end{equation*}
is a conformal mapping since it is a scaling and a rotation, both of which conserve the angle and direction. 
Let \(U\) be an open set in \(\Complex\) and \(\gamma,\eta : \clcl{a}{b} \to U\)  be two complex valued curves. Suppose that \(z_1 = \func{\gamma}{t_1} = \func{\eta}{t_2}\) then the angle between \(\gamma\) and \(\eta\) at \(z_1\) is defined as the angle between \(\func{\gamma'}{t_1}\) and \(\func{\eta'}{t_2}\), provided that they are nonvanishing. Let \(f: U \to \Complex\) be a holomorphic function then by chain rule 
\begin{equation*}
      \ODiff{\func{f}{\func{\gamma}{t}}}{t} = \func{f'}{\func{\gamma}{t}} \func{\gamma'}{t}
\end{equation*}
Viewing \(w = \func{f'}{\func{\gamma}{t}}\) as a \(\Reals^2\) linear transformation, we can deduce that \(f\) is conformal. 
\begin{align*}
      \ODiff{\func{f}{\func{\gamma}{t_1}}}{t} &= \func{f'}{z_1} \func{\gamma'}{t_1}\\
      \ODiff{\func{f}{\func{\eta}{t_2}}}{t} &= \func{f'}{z_1} \func{\eta'}{t_2}
\end{align*}
Thus given that \(\func{f'}{z_1} \neq 0\) the angle between \(\gamma\) and \(\eta\) at \(z_1\) is the same as \(f \circ \gamma\) and \(f \circ \eta\).